---
title: "Predictions of optimal pollinator visit rates to crops using simple mechanistic models"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Ignasi Bartomeus, Maxime Eeaerts, Mike Garratt, Jake Bishop, Tom Breeze?

\[Question: Should we invite Liam Kendall, given we use his data? Or is it fine like this, as is publicly available?\]

## Abstract

While the positive effect of pollinators on crop production has been clearly demonstrated, crop level predictions of how pollinator visit rates increase crops reproductive success remains elusive and requires time consuming experiments. Defining the functional form of this relationships is of pivotal importance to understand when pollination is limiting, and to develop operational visitation rate values that can be used by practitioners to rapidly assess pollination levels. Given the large variety of crops, and varieties, running detailed experiments for all of them is unlikely to be achieved in the short term. Here we propose and test a generalized mechanistic model based on biological first principles to infer optimal pollinator visitation rates for multiple crops. First we show that a non-inflicted functional form fits well empirical data. Second, we develop and test a flexible model to produce this functional forms from easy to obtain plant and pollinator traits such as plant's selfing rates, ovule number, and pollinator single visit efficiency. Finally, we apply our framework to 26 crops and ilustrate its flexibility using blueberries as a case study to show the advantages and limitations of the framework. An R package is provided to facilitate the uptake of this modelling framework by practitioners.

## Introduction

Agronomic research have a long history of quantifying the optimal inputs needed for optimizing crop production. For example, control trials have provided clear guidelines on optimal fertilizer use, irrigation, or pesticide application (ref?). However, insect pollination remained hard to measure, and was traditionally assumed to be either non-limiting, or easily covered by stocking managed pollinators *ad libitum* (Garibaldi et al 2020). However, in the last 20 years the importance of both managed and wild pollinators has been made evident, mainly trough synthesis papers showing a positive relationship between pollinator visitation and crop production (Garibaldi et al. 2013, Rader et al. 2016, Reilly et al. 2025). Consequently, the interest for managing crop pollination with both managed and wild pollinators has increased (Garibaldi et al 2015 Frontiers). Despite this efforts, the intrinsic variability across crops and varieties (Reilly et al. 2025), the assumption of linear relationships between pollinator visitation and yield efefcts (Eeaerts et al. 2026), and the confounding factors affecting yield beyond pollinators (Deguines et al. 2010), has prevented the emergence of clear guidelines suggesting optimal pollinator visitation rates for different crops (Garibaldi et al 2020).

Crop yield is affected by many factors, including fertilizers, water availability, or cropping density. Hence synthesis of observational studies risk confounding the effects of pollination on crops when pollinator availability is correlated with any of this factors (Chabert et al. 2026, Aizen et al 2025 PCI). Moreover, most analysis make statistical simplifications such as assuming a linear relationship, despite it has been shown that non linear, quadratic, relationship often fits better the data (Morris et al 2004, Eeaerts et al. 2026). The consequences is that even simple metrics such as how many visits per flower are needed to ensure pollination is not limiting are hard to extract from this type of observational analysis, even when crops are analyzed one by one (Reilly et al. 2025, Eeaerts et al. XXXX). As an alternative, controlled experiments, where crop resources are kept constant, and the number of visits per flower are tracked can provide detailed information on how many visits by different pollinators are needed to maximize crop yield (Refs Garratt, etc...). However, this experiments are terribly time consuming, and hence have only been done for a handful of crops (refs). Even within those crops, striking differences are observed depending of the cultivar, environmental conditions used, and the pollinator community available. For example, \[search for an example of a crop where recommended visitation rate changes by study - Maxime?\]. Overall, this suggests that experiments might be hard to generalize.

Here, we propose a well known functional form relating pollinator visitation to crop production: the non-inflicted curve (Morris et al. 2004). The advantage of this functional form over statistical quadratic fits is that its parameters have biological meaning. The curve is defined by three parameters which govern the relationship between pollinator visitation rates and crop seed set. Parameter $a$ sets the y-intercept (i.e., seed set for unvisited flowers), parameter $b$ sets the asymptote (i.e., maximum seed set attainable by a single flower), and parameter $c$ governs the rate of approach to the asymptote (Fig. 1). All this parameters can be estimated from experimental or observational data, but can be also derived from first principles. For example, existing databases on selfing rates can inform about the expected intercepts (Kleijn et al 2007, Siopa et al 20XX). Similarly, the number of fertile ovules can be obtained from botanic manuals informing the maximum seed set attainable per flower. Finally, the rate of approach can be derived from the pollinator efficiency in a per visit unit (ref on SVE). This opens the door to derive mechanistic relationships for a wide variety of crops and contexts.

Defining a clear framework on how to assess the pollinators contribution to yield will facilitate farmers and agronomists to incorporate rigurous assessments of pollination needs and pollinator management. Here, we provide functions to calculate optimal visitaion rates from experimental data when available, and most importantly, from easy to obtain biological parameters. By using a generalized mechanistic model we provide tools to assess any given crop in in any specific context rapidly. To facilitate the uptake of this tools, an R package accompains this article, along ilustrative case studies showing its flexibility and limitations.

## Methods

#### Modelling framework.

We use a functional form proposed by Morris et al. (2004) that describes a seed set model that accounts for two different ways a plant can produce seeds: through autonomous self-pollination (selfing) and through pollinator-assisted fertilization (outcrossing). In mathematical notation, the equation looks like this:

`seed set = ((b*a)/100) + (b-((b*a)/100)) * (1-exp(-c*visits))`

\[Note: I will use proper mathematical notation through the paper, for now, is R code\]

The formula is split into two main parts: a fixed baseline and a variable growth curve. The baseline (i.e. the intercept, representing autonomous seed set production) is defined by $b$ , the maximum potential seed set, and $a$, the percentage of that maximum that can be achieved through self-pollination (e.g., if $a=20$, the plant can produce 20% of its seeds without insect pollination). The pollinator contribution (i.e. the variable growth curve) represents the remaining seed potential that requires pollinators. Here, $c$ goberns the exponential saturation curve. It starts at $a$ and climbs toward $b$ as the number of visits increases. A higher $c$ means the plant reaches its maximum seed set with very few pollinator visits. This model show that there is a diminishing return on subsequent pollinator visits. Note that If $visits = 0$, the entire second half of the equation becomes 0 and only with the baseline counts. As $visits$ increases, the seed set grows quickly at first, then slows down as it approaches the maximum limit ($b$).

Optimal visitation rate (*OVR*) can be calculated as the visits required to achieve 95% (or any other chosen threshold) of potential seed set.

#### Data

We collected from the literature data on five studies measuring, for individual flowers, the seed set produced as a function of the number of visits received. This five studies cover different crop types (apple -REF-, strawberry -REF-, highbush blueberry -REF-, watermelon -REF- and lima bean -REF-). On this data we fitted three different models and compare them by AIC and R2. The three models are a simple linear relationship, a quadratic relationship and the non-inflicted functional form described above. Models were run using `lm()` and `nls()` functions in basic R.

#### Generalized model

As this data is rare in the literature, and hard to obtain experimentally, we derivd a generalized version of our modelling framework that rely on first principles and does not require hard won data. To that end, we set $b$ as the maximum seed set a flower can attain. This can be extracted from the crop basic biology, as it will match the per flower ovule production, a parameter often easy to obtain from the literature or to count in the field. Parameter $a$ , given in % of seedset without pollinators, can be approximated knowing plants selfing rates. That is, who many seeds the flower produce under prevented animal pollination. This information is available in the literature (e.g. Klein et al. 2007, Siopa et al 20XX) or can be calculated with basic low-effort experiments for your target cultivar. The trickier parameter to calculate is $c$. For this, we rely on pollinator per visit efficiency. Many studies measure seed set after a single visit of a pollinator. This values can be used to extrapolate how many visits a flower requires until fertilizing all its ovules. The average pollinator efficiency (SVE) can be calculated across the available pollinator pool, which is the representative metric, or calculated for different pollinator groups (see case study below for an example). This data is already available for many crops (e.g. Garratt, Page, etc...), and is expected to have lower variability across pollinators, than pollinator visitation rates (Vazquez et al. 2005 and Kleijn et al 2015). Obtaining this data from field observations requires only a moderate effort (REF on SVE protocol). After this first visit, it is known that each subsequent visit is less efficient (Kendall et al. 20XX), hence we can penalize each consecutive visit by a diminishing returns function. This penalization can be informed biologically, as we expect this diminishing returns should approach $b$. Hence, we define the diminishing returns function as `efficiency = SVE * (1 - (SVE / target_gap))^i` , where the `target_gap` is the difference between the baseline and the maximum attainable seed set, and $i$ is the number of visits after the first visit. When $i = 0$ (first visit), efficiency is equal to the single visit efficiency (SVE), but this value is exponentially lower at each subsequent visit.

To test the flexibility of this approach, we first compiled information on $a$, $b$ and $SVE$ parameters for 26 major crops by scanning the literature. Note that those are informative generic values, but we encourage users to look up for values for their crop variety, an region. Next, we visually compare this data for the five crops for which we have detailed data.

Finally, we delve into the flexibility of the approach in a case study on blueberry pollination (Kendall et al. 20XX). This allow us to show how to customize the values for different cultivars or pollinator communities.

```{r}
#| echo: false
#| warning: false

#Source functions. Should be made into a package.
source(file = "scripts/functions.R")

c <- calculate_visits(a = 20, b = 100, SVD = 40, to_ = 15, 
                 plot_ = TRUE, col_ = 2)
ovr <- calculate_required_visits(a = 20, b = 100, c = c$c_parameter)
abline(h = 100, lty = 2, col = "brown")
lines(y = c(-2,95), x = c(ovr,ovr), lty = 2, col = "blue")
lines(y = c(-2,40), x = c(1,1), lty = 2)
points(x = 1, y = 40, pch = 19)
text(x = 2, y = 20, "a = selfing rate", col = "darkgrey", cex = 0.7)
text(x = 4.3, y = 40, "SVE = single visit efficiency", cex = 0.7)
text(x = 6, y = 58, "c = diminishing returns param.", col = "red", cex = 0.7)
text(x = 2, y = 95, "b = maximum seed set", col = "brown", cex = 0.7)
text(x = 12, y = 80, "OVR = optimal visits", col = "blue", cex = 0.7)
```

**Figure 1:** Theoretical explanation of the modelling framework. The plot shows the intercept value defined as $a$ (in grey), which is defined as the percentage of seeds obtained with no animal pollination or selfing rates. Parameter $b$ defines the maximum seed set attainable seed set (in brown). Parameter $c$ describes the form of the curve (in red). We also show how single visit efficiency (SVE, in black) can be used to fit the curve, and the optimal visitation rate (OVR, in blue) defined as the number of visits required to attain the 95% of the maximum seed set.

## Results

The non-inflicted functional form show a better or equal performance than the quadratic or linear fit for all five crops (Fig. 2), demonstrating that is a useful modelling framework. Despite it needs to estimate three parameters ($a$, $b$, and $c$) agaist two parameters for the lineal model, AIC values are better (DeltaAIC \> 2, or similar DeltaAIC \~ 0). Most importantly, when visually inspecting the fitted lines, our model avoid the unrealistic linear increase of seed set beyond maximum number of ovules ($b$), and the reduction of seed set at high visitation rate shown by some hump-shaped quadratic fits. Note that despite this seed set reduction has been shown in some crops (Saez et al XXX), the mechanisms are often different (e.g. stigma damage). This potential negative effects of higer than needed visiaion rates can be assessed by flagging visitation rates values that are well beyond the calculated optimal.

Overall, we highlight the large variability observed in all models and crops, indicating that even on controlled experiments, there is a wide uncertainty in the effects of pollinator visitation.

```{r}
#| echo: false
#| warning: false

#read observational data
dat <- read.csv("data/VisitationRates_Data.csv")
#make a dataset per crop
#unique(dat$Crop)
LimaBean <- subset(dat, Crop == "Lima.bean")
Strawberry <- subset(dat, Crop == "Strawberry") 
SHighbushBlueberry <- subset(dat, Crop == "Southern.Highbush.Blueberry")  
NHighbushBlueberry <- subset(dat, Crop == "Northern.Highbush.Blueberry")
Rabbiteye.Blueberry <- subset(dat, Crop == "Rabbiteye.Blueberry")
Apple <- subset(dat, Crop == "Apple")
Watermelon <- subset(dat, Crop == "Watermelon")
```

```{r}
#| echo: false
#| warning: false


#AIC and R2 in Real_crops.qmd

#apple
#fit models
model_nls <- fit_data(seedset = Apple$PollinationValue, visitation = Apple$Number.of.visits, simplify = "model")
model_lm <- lm(Apple$PollinationValue ~ Apple$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Apple)

#Extract a, b and c values
mf <- fit_data(seedset = Apple$PollinationValue, visitation = Apple$Number.of.visits, simplify = "params")

#Plot
par(mfrow=c(2,3))
plot(Apple$PollinationValue ~ jitter(Apple$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey", main = "Apple")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), add = TRUE, col = "#009E73")

#Strawberry
Strawberry <- subset(Strawberry, Pollinator %in% c("Apis mellifera", "Andrena.spp", "Ceratina.spp"))

#fit models
model_nls <- fit_data(seedset = Strawberry$PollinationValue, visitation = Strawberry$Number.of.visits, simplify = "model",
                      a_start = 10, b_start = 100, c_start = 0.5)
model_lm <- lm(Strawberry$PollinationValue ~ Strawberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Strawberry)

#Extract a, b and c values
mf <- fit_data(seedset = Strawberry$PollinationValue, visitation = Strawberry$Number.of.visits, simplify = "params", a_start = 10, b_start = 100, c_start = 0.5)

#Plot
plot(Strawberry$PollinationValue ~ jitter(Strawberry$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey", xlim = c(0,20), , main = "Strawberry")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = 30, add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")

#Lime Bean
#fit models
model_nls <- fit_data(seedset = LimaBean$PollinationValue, visitation = LimaBean$Number.of.visits, simplify = "model")
model_lm <- lm(LimaBean$PollinationValue ~ LimaBean$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = LimaBean)

#Extract a, b and c values
mf <- fit_data(seedset = LimaBean$PollinationValue, visitation = LimaBean$Number.of.visits, simplify = "params")

#Plot
plot(LimaBean$PollinationValue ~ jitter(LimaBean$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey", main = "Lime Bean")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), add = TRUE, col = "#009E73")

#Blueberry
SHighbushBlueberry <- subset(SHighbushBlueberry, PollinationMetric == "Fruit.weight")

#fit models
model_nls <- fit_data(seedset = SHighbushBlueberry$PollinationValue, 
                      visitation = SHighbushBlueberry$Number.of.visits, simplify = "model")
model_lm <- lm(SHighbushBlueberry$PollinationValue ~ SHighbushBlueberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = SHighbushBlueberry)

#Extract a, b and c values
mf <- fit_data(seedset = SHighbushBlueberry$PollinationValue, 
               visitation = SHighbushBlueberry$Number.of.visits, simplify = "params")

#Plot
plot(SHighbushBlueberry$PollinationValue ~ jitter(SHighbushBlueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", col = "grey", main = "Blueberry")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(SHighbushBlueberry$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), add = TRUE, col = "#009E73")

#Watermelon
#fit models
model_nls <- fit_data(seedset = Watermelon$PollinationValue, 
                      visitation = Watermelon$Number.of.visits, simplify = "model", a_start = 75, b_start = 300)
model_lm <- lm(Watermelon$PollinationValue ~ Watermelon$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Watermelon)

#Extract a, b and c values
mf <- fit_data(seedset = Watermelon$PollinationValue, 
               visitation = Watermelon$Number.of.visits, simplify = "params", 
               a_start = 75, b_start = 300)

#Plot
plot(Watermelon$PollinationValue ~ jitter(Watermelon$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Seed set", col = "grey", main = "Watermelon")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(Watermelon$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), add = TRUE, col = "#009E73")
#Add legend with colors.
par(mfrow=c(1,1))

```

**Figure 2:** Seed set as a function of visitation rate for data collected on five different crops. a) apple, b) strawberry, c) highbush blueberry, d) watermelon, e) lima bean. Blue lines repsent the linear fit, green lines the quadratic fits and orange lines the non-inflicted fucntional form.

\[Note: AIC and R2 values to appendix? or in text?\]

\[**Question**: This data is tricky. Apple data from Mike is ok. Strawberry data does not saturate, which I wonder if they cover the right visits range (maybe we can add the open treatment to fix the upper bound? Maxime?. Lima Bean is terrible... zero is the most obtained value for any visitation rate!! Other studies on Lima bean show high autocompatiblity... should we discard this one? Blueberry looks good, but we use weight. Maxime, how would you explain without complicating the story that weight and seed set is highly correlated in methods? Watermelon is also ok-ish, but the V = 30 is the open visitation treatment... so I do not know this value for sure (not in the paper, I think), and this will affect the shape a bit... BUT overall, variation is huge eveywhere. Ideas on how to make this section more powerful are welcome\]

Using the generalized modelling approach, we calculated optimal visitation rates for 26 crops (Fig. 3). Optimal visitation rates range from \~25 for blueberry, kiwi, peppers or squash, to a \< 5 visits for pears or Lima beans.

\[Note: When plotting OVR as a function of a, b and SVE parameters, we see how it emerges from a combination of those, and is not strongly influenced by any single parameter. Should we show that? Or is not relevant? Or expected?\]

```{r}
#| echo: false
#| warning: false

#read data
dat <- read.csv("data/VisitationRates_ParameterData.csv")
#head(dat)
dat <- dat[-24,] #removing cheak peas, which is wierd as a = b = SV

#param a is in %
dat$a <- (dat$a / dat$b)*100

#SV we can take the average
dat$SV <- rowMeans(dat[,c("Honeybee", "Bumblebee", 
                          "Other.wild.bee", "Hoverfly", "Other")], na.rm = TRUE)
#But some are wierd
dat$SV[which(dat$Crop == "cherry")] <- 0.25 #Using only wild bees.
dat$SV[which(dat$Crop == "strawberry")] <- 150 #Using max efficiency...

#output data frame
out <- data.frame(crop = rep(NA,nrow(dat)), 
                  family = rep(NA,nrow(dat)),
                  a = rep(NA,nrow(dat)), 
                  b = rep(NA,nrow(dat)), 
                  c = rep(NA,nrow(dat)), 
                  SVE = rep(NA,nrow(dat)), 
                  OVR = rep(NA,nrow(dat)))
#loop it
for(i in 1:nrow(dat)){
  #extract c
  params <- calculate_visits(a = dat$a[i], b = dat$b[i], SVD = dat$SV[i], 
                             plot_ = FALSE)
  #warning can be dissmissed.
  #And visits needed.
  OVR_model <- calculate_required_visits(a = dat$a[i], b = dat$b[i], c_val = params$c_parameter) 
  #store the data
  out$crop[i] <- dat$Crop[i]
  out$family[i] <- dat$family[i]
  out$a[i] <- dat$a[i]
  out$b[i] <- dat$b[i]
  out$c[i] <- params$c_parameter
  out$SVE[i] <- dat$SV[i]
  out$OVR[i] <- OVR_model
}

par(mfrow=c(1,2), mar=c(8, 4, 1.5, 0.5))
bymedian <- with(out, reorder(family, -OVR, median))
boxplot(out$OVR ~ bymedian, las = 2, xlab = "", ylab = "OVR")
bymedian <- with(out, reorder(crop, -OVR, median))
boxplot(out$OVR ~ bymedian, las = 2, xlab = "", ylab = "OVR",
        cex.axis = 0.5)
par(mfrow=c(1,1), mar=c(4, 4, 1.5, 0.5))

#parameter influence? To SUp Mat?
#scatter.smooth(out$OVR ~ out$a, las = 1)
#scatter.smooth(out$OVR ~ out$c, las = 1)
#scatter.smooth(out$OVR ~ out$SVE, las = 1)
#It's a combination, but c is equivalent to OVR 
```

**Figure 3:** Model predictions of optimal visitation rates for 26 main EU crops grouped by plant families (A) and for individual crops (B). Boxplots show medians and ...

\[**Question**: numbers look a bit high, and some even odd. I think we can use Siopa for a in % and try again. b is a bit high in general, as (maybe) ovules \> max seed set. SVD is what we can obtain... maybe using it in % of max seed set woul help standardize for varieties? We can discuss this in person\].

If we focus on the five crops for which we have also empiricall data, we can visually compare the models parameters for the data-driven values and mechanistic values (Fig. 4). In general, there is a decent agreement, but we tend to overestimate $b$ . This is not suprising, given the large variability observed in empirical data, where even well pollinated flowers show a large variation in seed set. As a consequence, OVR tend to be higher in the mechanistic model than in the empiriall data. Note that for example, watermelon seems to require less visits under the mechanistic model. This is because single visit values in the literature (Ref used, but also see REF), are much higher than for the data used to fit the model (\~200 seeds per visit vs \~50 seeds per visit). When using the 50 seeds per visit value the discrepancy is reduced. This indicates the importance of choosing the initial parameters well.

```{r}
#| echo: false
#| warning: false

#load params
all_data <- read.csv("data/parameters_real_crops.csv")
#head(all_data)
all_data <- all_data[-c(5,6),] #removing BB varieties

par(mfrow=c(2,2), mar=c(3, 4, 1, 0.5))
plot(all_data$OVR, all_data$OVR_model, 
     ylim = c(0,max(all_data$OVR,all_data$OVR_model)+5), 
     xlim = c(0,max(all_data$OVR,all_data$OVR_model)+5),   
     xlab = "Obs_OVR", ylab = "Model_OVR", 
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)
#Not too satisfactory... specially strawberry, but also watermelon and lima bean
plot(all_data$a, all_data$a_model, 
     ylim = c(0,max(all_data$a,all_data$a_model)+5), 
     xlim = c(0,max(all_data$a,all_data$a_model)+5),  
     xlab = "Obs_a", ylab = "Model_a",
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)

plot(all_data$b, all_data$b_model, 
     ylim = c(0,max(all_data$b,all_data$b_model)+5), 
     xlim = c(0,max(all_data$b,all_data$b_model)+5), 
     xlab = "Obs_b", ylab = "Model_b",
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)

plot(all_data$c, all_data$c_model, 
     ylim = c(0,max(all_data$c,all_data$c_model)+5), 
     xlim = c(0,max(all_data$c,all_data$c_model)+5),  
     xlab = "Obs_c", ylab = "Model_c",
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)
par(mfrow=c(1,1), mar=c(4, 4, 1.5, 0.5))
```

**Figure 4:** Comparision for the four main parameters ($a$, $b$, $c$, and $OVR$) for data driven and mechanistic models. Using the same five crops, we compare empirically derived $a$, $b$, $c$, and $OVR$ parameters, with $a$, $b$, $c$, and $OVR$ predictions in a scatter plot with the 1:1 line plotted.

\[**Question:** I am unsure if this is a fair comparison. Empirical data is not perfect, and b is always lower due to variability. I wonder if using OVR at 80% instead than at 95% can fix this. I think we can check Siopa for a (maybe is better as it will be in %). Maxime, do you have this data already processed?\]

We showed that with basic data, we can generate plausible functional forms for many crops, and estimate their expected optimal visitation rates. However, as expected those are at quite coarse resolution. Our framework allow refining this predictions if the right data is available. Using detailed data collected by Kendall et al. 2024 for blueberry, we first predict the relationship between visitation and fruit weight for different varieties (Fig. 5A), and second for different pollinator communities (Fig, 5B). Note that in blueberry, fruit weight and seed set are highly correlated (REF?), and allow us to use it under this same framework, but for most crops fruit weight might be decoupled from seed set and this trick might not work \[**Question**: can we transform weight to seed set and use seed set in figures to avoid confunding readers? Maxime?\]. We see that detailed information on selfing rates and maximum yield attainable change the predictions for the different varieties. Similarly, different pollinator communities might affect the number of visits needed to attain optimal fruit production. Specifically, for Southern Highbush blueberry, receiving only honeybee visits require more visits than recieving only*Tetragonula carbonaria* vists, with mixed communities showing an intermediate result.

```{r}
#| echo: false
#| warning: false

#read data
dat <- read.csv("data/VisitationRates_Data.csv")
#make a dataset per crop
SHighbushBlueberry <- subset(dat, Crop == "Southern.Highbush.Blueberry")  
SHighbushBlueberry <- subset(SHighbushBlueberry, PollinationMetric == "Fruit.weight")
NHighbushBlueberry <- subset(dat, Crop == "Northern.Highbush.Blueberry")
NHighbushBlueberry <- subset(NHighbushBlueberry, PollinationMetric == "Fruit.weight")
Rabbiteye.Blueberry <- subset(dat, Crop == "Rabbiteye.Blueberry")
Rabbiteye.Blueberry <- subset(Rabbiteye.Blueberry, PollinationMetric == "Fruit.weight")

#We can use quantiles, but mean of 0 and and max visits, or simply min and max, might also work.

#For SHB
SHB_a_real <- quantile(subset(SHighbushBlueberry, 
                              Number.of.visits == 0)$PollinationValue)[2]
#max(SHighbushBlueberry$Number.of.visits)
SHB_b <- quantile(subset(SHighbushBlueberry, 
                         Number.of.visits == 15)$PollinationValue)[4]
SHB_SVE <- mean(subset(SHighbushBlueberry, 
                       Number.of.visits == 1)$PollinationValue)
#We need a in %
SHB_a <- (SHB_a_real/ SHB_b)*100

#For NHB
NHB_a_real <- quantile(subset(NHighbushBlueberry, 
                              Number.of.visits == 0)$PollinationValue)[2]
#max(NHighbushBlueberry$Number.of.visits)
NHB_b <- quantile(subset(NHighbushBlueberry, 
                         Number.of.visits == 15)$PollinationValue)[4]
NHB_SVE <- mean(subset(NHighbushBlueberry, 
                       Number.of.visits == 1)$PollinationValue)
#We need a in %
NHB_a <- (NHB_a_real/ NHB_b)*100

#For REB
REB_a_real <- quantile(subset(Rabbiteye.Blueberry, 
                              Number.of.visits == 0)$PollinationValue)[2]
#max(Rabbiteye.Blueberry$Number.of.visits)
REB_b <- quantile(subset(Rabbiteye.Blueberry, 
                         Number.of.visits == 10)$PollinationValue)[4]
REB_SVE <- mean(subset(Rabbiteye.Blueberry, 
                       Number.of.visits == 1)$PollinationValue)
#We need a in %
REB_a <- (REB_a_real/ REB_b)*100

#calculate param c
SHB_c <- calculate_visits(a = SHB_a, b = SHB_b, SVD = SHB_SVE, to_ = 15)$c_parameter
NHB_c <- calculate_visits(a = NHB_a, b = NHB_b, SVD = NHB_SVE, to_ = 15)$c_parameter
REB_c <- calculate_visits(a = REB_a, b = REB_b, SVD = REB_SVE, to_ = 15)$c_parameter

#now we can plot all three
par(mfrow=c(1,2))
plot(SHighbushBlueberry$PollinationValue ~ jitter(SHighbushBlueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", type = "n", main = "Different varieties",
     ylim = c(0,4), xlim = c(0,15))
plot_visits(SHB_a, SHB_b, SHB_c, to_ = 15, 
            add_ = TRUE, col = "#223366")
plot_visits(NHB_a, NHB_b, NHB_c, to_ = 15, 
            add_ = TRUE, col = "#A3B18A")
plot_visits(REB_a, REB_b, REB_c, to_ = 15, 
            add_ = TRUE, col = "#B9D1EA")

#Now different pollinator compositions.
#For SHB
SHB_a_real <- quantile(subset(SHighbushBlueberry, 
                              Number.of.visits == 0)$PollinationValue)[2]
#We need a in %
SHB_b <- quantile(subset(SHighbushBlueberry, 
                         Number.of.visits == 10)$PollinationValue)[4]
SHB_a <- (SHB_a_real/ SHB_b)*100

#pollinators available
#unique(SHighbushBlueberry$Pollinator)
SHBmix <- subset(SHighbushBlueberry, Pollinator == "pollinator.mix")
SHBhb <- subset(SHighbushBlueberry, Pollinator == "Apis.mellifera")
SHBbb <- subset(SHighbushBlueberry, Pollinator == "Tetragonula.carbonaria")

#calculate different SVD values
#SHB_SVE_mix <- mean(subset(SHBmix, 
#                           Number.of.visits == 2)$PollinationValue)/2 #diviing by 2 penalizes a lot... see below
SHB_SVE_hb <- mean(subset(SHBhb, 
                          Number.of.visits == 1)$PollinationValue)
SHB_SVE_bb <- mean(subset(SHBbb, 
                          Number.of.visits == 1)$PollinationValue)
SHB_SVE_mix <- (SHB_SVE_hb + SHB_SVE_bb)/2


#calculate param c
SHB_c_mix <- calculate_visits(a = SHB_a, b = SHB_b, SVD = SHB_SVE_mix, to_ = 15)$c_parameter
SHB_c_hb <- calculate_visits(a = SHB_a, b = SHB_b, SVD = SHB_SVE_hb, to_ = 15)$c_parameter
SHB_c_bb <- calculate_visits(a = SHB_a, b = SHB_b, SVD = SHB_SVE_bb, to_ = 15)$c_parameter
#now we can plot all three
plot(SHighbushBlueberry$PollinationValue ~ jitter(SHighbushBlueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", type = "n", main = "Pollinator composition",
     ylim = c(0,2), xlim = c(0,10))
plot_visits(SHB_a, SHB_b, SHB_c_mix, to_ = 10, 
            add_ = TRUE, col = "#D55E00")
plot_visits(SHB_a, SHB_b, SHB_c_hb, to_ = 10, 
            add_ = TRUE, col = "#E69F00")
plot_visits(SHB_a, SHB_b, SHB_c_bb, to_ = 10, 
            add_ = TRUE, col = "darkred")

par(mfrow=c(1,1))
```

**Figure 5:** We use blueberry as a case study to ilustrate how to go from generic predictions to tailored predictions to your context. Using blueberry data from Kendall et al 2024, we show A) that we can predict curves for different varieties (Northern, Southern and Rabiteye), and that B) different pollinator communities (Honeybees dominated, wild bee dominated, and mixed) in southern highbush blueberry show different curves.

## Discussion

We show how modelling the relationship between insect crop flower visitation rates and seed production is not only possible, but efficient and salable. By developing a simple mechanistic model based on ecological first principles, we presented generic optimal pollinator values for 26 common crops, and show how this values can be fine tuned to specific contexts. This is specially important given the high variability expected for this relationships depending on the cultivar of interest, and the available pollinator community.

A core assumption of of our modelling framework is that only pollinator-mediated effects are modeled. Rather than a limitation, we see this as a feature of the model, given it is interpreted correctly. When assessing pollinator limitation (ref?), our mechanistic model do not confound pollinator effects with confounding variables such as agricultural intensification which can mask effects on final yield (Eeaerts et al. In prep). Hence, rather than directly comparable, observational data and our model outputs can complement each other to discover if something else is limiting yield besides pollinator levels. Similarly, when making recommendations of optimal pollinator levels (Garibaldi et al 2020), the model correctly assume those are independent of other management practices. Note that when fitting visitation - seed set curves from observational data directly, confounding factors can still apply. In fact, this is precisely why linking seed set to yield has been elusive so far, and even well researched crops lack clear guidelines on which are the optimal pollinator levels.

Despite the clear framework we provide, the choice of the input parameters is still not trivial. Ideally, selfing rates, maximum attainable seed set and pollinator efficiency should be collected for your cultivar of interest and pollinator pool. When this is not possible, informed choices can provide good proxies, but given the large variability in some crops species among cultivars, caution is needed when selecting those parameters. As an example, we found that our model match the empirical data expectations for watermelon when we use information on selfing rates, maximum seed production and pollinator single visit efficiency from the same study (Di Trani et al. 2024), but not when using available single visit efficiency data from other studies (REFS). \[**Note**; Maybe we need a Figure 4 using same data for mechanistic and non mechanistic models in sup mat or main text?\]. While we predict \~5 visits are neede, while \~20 are needed for Di Trani et al. 2024 dataset, this is not a necessarely a bug of the model, as a value of \~5 visits might be sufficient in other contexts. Hence, the values provided for the 26 crops analyzed are generic values for this kind of crops, but we expect to vary substantially across cultivars and pollinator communities within crops. In fact, our case study using blueberry data confirms how fine-tuning the input parameters to specific contexts can change the outputs in meaningful ways.

While we focus on the pathway between visitation rate per flower and seed set per fruit, we think expanding the framework to cover the full cycle from pollinator abundance to yield is easily achievable. Most ecological datasets (Allen-Perkins et al 2024) measure pollinator activity as visits per meter and hour rather than in vistis per flower. This is because the most common method is walking a fixed length transect during a fixed time while annotating insects interacting with the plants flowers. Knowing the expected flowers per square meter for your field, flower longevity, and using simple conversion rates, we can easily convert visits per m2 and hour to visits per flower (implemented in our R package as `V_transect_to_flower(V_transect = 50, flw_x_m2 = 400, lifespan = 6).`

Similarly, while we work at the flower level, we should think on this flower as an average flower. Visitation rates of less then one visit per flower are hence relevant, as those imply not all flowers will get at least one visit, as can be the case in pollinator depauperate areas (ref?). Hence, to turn our seed set per flower into yield, we would (simply) need to know the relationship between seed set and fruit weight. This can be, again, obtained from the literature, or calculated empirically using a few selected fruits covering a large range of weights for which we also count seeds. If desired, this relationship can be truncated to account for abortions if the data suggest fruits with less than a minimum number of seeds tend to abort. Then, calculating yield implies: `Yield = Flowers/Ha * Visits/Flowers * SeedSet/Visits * FruitWeight/SeedSet = Weight/Ha.` Of course, the more parameters we estimate, the most uncertain is the final value obtained, and proper error propagation is needed (i.e. accounting for the error or variability in all parameter inputs).\
\
While we still lack high quality data on the input parameters for most cultivars, we hope that having a unified framework will speed up the collection of this data, and its addition to unified databases (e.g. CropPol, ...). In this paper we provide the first raw estimates of how many visits per flower we expect different crops to need to ensure pollen is not limiting. This value is critical for starting to measure pollination at the same level that other agricultural inputs such as fertilization or pesticides, and improve management beyond actual guesstimates (Garibaldi et al. 2020; refs). With this guidelines, visitation rates well below the optimal recommendations will require measures such as adding managed pollinators (ref... Benjamin?), or enhancing within field (e.g. flower stripes; Schepper et al. ) or out of field (e.g. landscape linear elements; REF) semi-natural elements to improve pollination (Garibaldi et al frontiers). On the contrary, values well above the optional recommendations indicate that managed pollinators might be at too high densities. Given the known negative effects of managed pollinators at high densities (REFS: Magrach), this can be a powerful tool to convince farmers to reduce its numbers when visitation rates are too high. In fact, many crops will suffer negative effects when managed pollinators are too abundant, as stigmas can be damaged (Saez et al) and having critical numbers on when high visitation rates are too high can rapidly become an useful tool to limit the number of managed pollinators in agro-ecosystems.

We encourage researchers to use our modelling framework, for which a dedicate R package is provided. Only by contrasting modeled and observed data we can enhance models and make then useful (Dietze et al). It's been more than five years since we proposed to better monitor pollinator inputs in agro-ecosystems (Garibaldi et al 2020). While this paper has received more than 100 citations, just a handful of those actually provide critical thresholds for pollinator optimal visitation rates. The time is ripe for developing this numbers and use them in applied settings.

## Acknowledgements

VALOR, AGRI4POL

## References

TBA
