---
title: "Real crops"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

## Crop selection

We will explore:

- Garratt Apple data

- Kendall Blueberry data

- Chirila Lima Bean data

- Kandori Strawberry data

- Di Trani Watermelon data

## Fit nls's

Load functions and data

```{r}
source(file = "scripts/functions.R")

dat <- read.csv("data/VisitationRates_Data.csv")
#make a dataset per crop
unique(dat$Crop)
LimaBean <- subset(dat, Crop == "Lima.bean")
Strawberry <- subset(dat, Crop == "Strawberry") 
SHighbushBlueberry <- subset(dat, Crop == "Southern.Highbush.Blueberry")  
NHighbushBlueberry <- subset(dat, Crop == "Northern.Highbush.Blueberry")
Rabbiteye.Blueberry <- subset(dat, Crop == "Rabbiteye.Blueberry")
Apple <- subset(dat, Crop == "Apple")
Watermelon <- subset(dat, Crop == "Watermelon")
```

### Apple

```{r}
#head(Apple)
#unique(Apple$Pollinator) #[1] "Bombus terrestris"    "Apis mellifera"       #"Episyrphus balteatus" "Osmia bicornis"      
#Apple <- subset(Apple, Pollinator %in% c("Apis mellifera", "Osmia bicornis", "Control"))

#fit models
model_nls <- fit_data(seedset = Apple$PollinationValue, visitation = Apple$Number.of.visits, simplify = "model")
model_lm <- lm(Apple$PollinationValue ~ Apple$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Apple)

#Test fit
AIC_nls <- AIC(model_nls) #2790
AIC_lm <- AIC(model_lm) #2834
AIC_cuadratic <- AIC(model_cuadratic) #2823

#Extract R2
R2_nls <- pseudeR2(model_nls, Apple$PollinationValue) #0.09
R2_lm <- summary(model_lm)$r.squared #0.01
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.03

#Extract a, b and c values
mf <- fit_data(seedset = Apple$PollinationValue, visitation = Apple$Number.of.visits, simplify = "params")

#Plot
plot(Apple$PollinationValue ~ jitter(Apple$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
#save values to compare with model. let's create a table
all_data <- data.frame(crop = c("apple"),
                       D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                       D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls,   
                       a = mf["a"],
                       b = mf["b"],
                       c = mf["c"],
                       OVR = calculate_required_visits(a = mf["a"],
                                                       b = mf["b"],
                                                       c_val = mf["c"]))
```

Best fit according to AIC is our nls model.

### Strawberry

```{r}
#head(Strawberry)
#table(Strawberry$Pollinator)  
#unique(Strawberry$PollinationMetric)
#needed?
Strawberry <- subset(Strawberry, Pollinator %in% c("Apis mellifera", "Andrena.spp", "Ceratina.spp"))

#fit models
model_nls <- fit_data(seedset = Strawberry$PollinationValue, visitation = Strawberry$Number.of.visits, simplify = "model",
                      a_start = 10, b_start = 100, c_start = 0.5)
model_lm <- lm(Strawberry$PollinationValue ~ Strawberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Strawberry)

#Test fit
AIC_nls <- AIC(model_nls) #6784
AIC_lm <- AIC(model_lm) #6783
AIC_cuadratic <- AIC(model_cuadratic) #6784

#Extract R2
R2_nls <- pseudeR2(model_nls, Strawberry$PollinationValue) #0.1
R2_lm <- summary(model_lm)$r.squared #0.09 #again best!
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.1

#Extract a, b and c values
mf <- fit_data(seedset = Strawberry$PollinationValue, visitation = Strawberry$Number.of.visits, simplify = "params", a_start = 10, b_start = 100, c_start = 0.5)

#Plot
plot(Strawberry$PollinationValue ~ jitter(Strawberry$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey", xlim = c(0,30))
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = 30, add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
#save values to compare with model. let's create a table
sb_data <- data.frame(crop = c("strawberry"),
                      D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                      D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls,  
                       a = mf["a"],
                       b = mf["b"],
                       c = mf["c"],
                       OVR = calculate_required_visits(a = mf["a"],
                                                       b = mf["b"],
                                                       c_val = mf["c"]))
all_data <- rbind(all_data, sb_data)
```

All models perform equally bad... but ours is still more realistic.

## Lima Bean

```{r}
#head(LimaBean)
#unique(LimaBean$Pollinator) #"Apis mellifera"    
#unique(LimaBean$PollinationMetric) #"Apis mellifera"    

#fit models
model_nls <- fit_data(seedset = LimaBean$PollinationValue, visitation = LimaBean$Number.of.visits, simplify = "model")
model_lm <- lm(LimaBean$PollinationValue ~ LimaBean$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = LimaBean)

#Test fit
AIC_nls <- AIC(model_nls) #608
AIC_lm <- AIC(model_lm) #606 #BEST!!
AIC_cuadratic <- AIC(model_cuadratic) #608

#Extract R2
R2_nls <- pseudeR2(model_nls, LimaBean$PollinationValue) #0.1
R2_lm <- summary(model_lm)$r.squared #0.09 #again best!
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.1

#Extract a, b and c values
mf <- fit_data(seedset = LimaBean$PollinationValue, visitation = LimaBean$Number.of.visits, simplify = "params")

#Plot
plot(LimaBean$PollinationValue ~ jitter(LimaBean$Number.of.visits), las = 1, xlab = "Visits", ylab = "Seed set", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(LimaBean$Number.of.visits), add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
lb_data <- data.frame(crop = c("lima bean"),
                      D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                      D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls,  
                      a = mf["a"],
                      b = mf["b"],
                      c = mf["c"],
                      OVR = calculate_required_visits(a = mf["a"],
                                                      b = mf["b"],
                                                      c_val = mf["c"]))
all_data <- rbind(all_data, lb_data)
```

All models perform equally bad...

### Southern Highbush Blueberry

```{r}
#head(SHighbushBlueberry)
#unique(SHighbushBlueberry$Pollinator) 
#unique(SHighbushBlueberry$PollinationMetric)  

SHighbushBlueberry <- subset(SHighbushBlueberry, PollinationMetric == "Fruit.weight")

#fit models
model_nls <- fit_data(seedset = SHighbushBlueberry$PollinationValue, 
                      visitation = SHighbushBlueberry$Number.of.visits, simplify = "model")
model_lm <- lm(SHighbushBlueberry$PollinationValue ~ SHighbushBlueberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = SHighbushBlueberry)

#Test fit
AIC_nls <- AIC(model_nls) #176
AIC_lm <- AIC(model_lm) #175 #BEST!!
AIC_cuadratic <- AIC(model_cuadratic) #176

#Extract R2
R2_nls <- pseudeR2(model_nls, SHighbushBlueberry$PollinationValue) #0.1
R2_lm <- summary(model_lm)$r.squared #0.09 #again best!
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.1

#Extract a, b and c values
mf <- fit_data(seedset = SHighbushBlueberry$PollinationValue, 
               visitation = SHighbushBlueberry$Number.of.visits, simplify = "params")

#Plot
plot(SHighbushBlueberry$PollinationValue ~ jitter(SHighbushBlueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(SHighbushBlueberry$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
shbb_data <- data.frame(crop = c("SHighbushBlueberry"),
                        D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                        D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls, 
                      a = mf["a"],
                      b = mf["b"],
                      c = mf["c"],
                      OVR = calculate_required_visits(a = mf["a"],
                                                      b = mf["b"],
                                                      c_val = mf["c"]))
all_data <- rbind(all_data, shbb_data)
```

Quite similar fits again...

### Northern Highbush Blueberry

```{r}
#head(NHighbushBlueberry)
#unique(NHighbushBlueberry$Pollinator) 
#unique(NHighbushBlueberry$PollinationMetric) 

NHighbushBlueberry <- subset(NHighbushBlueberry, PollinationMetric == "Fruit.weight")

#fit models
model_nls <- fit_data(seedset = NHighbushBlueberry$PollinationValue, 
                      visitation = NHighbushBlueberry$Number.of.visits, simplify = "model")
model_lm <- lm(NHighbushBlueberry$PollinationValue ~ NHighbushBlueberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = NHighbushBlueberry)

#Test fit
AIC_nls <- AIC(model_nls) #937
AIC_lm <- AIC(model_lm) #939
AIC_cuadratic <- AIC(model_cuadratic) #935 ##BEST!

#Extract R2
R2_nls <- pseudeR2(model_nls, NHighbushBlueberry$PollinationValue) #0.1
R2_lm <- summary(model_lm)$r.squared #0.09 #again best!
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.1

#Extract a, b and c values
mf <- fit_data(seedset = NHighbushBlueberry$PollinationValue, 
               visitation = NHighbushBlueberry$Number.of.visits, simplify = "params")

#Plot
plot(NHighbushBlueberry$PollinationValue ~ jitter(NHighbushBlueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(NHighbushBlueberry$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
nhbb_data <- data.frame(crop = c("NHighbushBlueberry"),
                        D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                        D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls, 
                        a = mf["a"],
                        b = mf["b"],
                        c = mf["c"],
                        OVR = calculate_required_visits(a = mf["a"],
                                                        b = mf["b"],
                                                        c_val = mf["c"]))
all_data <- rbind(all_data, nhbb_data)
```

Nicer?

### Rabbiteye Blueberry

```{r}
#head(Rabbiteye.Blueberry)
#unique(Rabbiteye.Blueberry$Pollinator) 
#unique(Rabbiteye.Blueberry$PollinationMetric) 

Rabbiteye.Blueberry <- subset(Rabbiteye.Blueberry, PollinationMetric == "Fruit.weight")

#fit models
model_nls <- fit_data(seedset = Rabbiteye.Blueberry$PollinationValue, 
                      visitation = Rabbiteye.Blueberry$Number.of.visits, simplify = "model")
model_lm <- lm(Rabbiteye.Blueberry$PollinationValue ~ Rabbiteye.Blueberry$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Rabbiteye.Blueberry)

#Test fit
AIC_nls <- AIC(model_nls) #20
AIC_lm <- AIC(model_lm) #18
AIC_cuadratic <- AIC(model_cuadratic) #19

#Extract R2
pseudeR2(model_nls, Rabbiteye.Blueberry$PollinationValue) #0.02
summary(model_lm)$r.squared #0.02
summary(model_cuadratic)$r.squared #0.02

#Extract a, b and c values
mf <- fit_data(seedset = Rabbiteye.Blueberry$PollinationValue, 
               visitation = Rabbiteye.Blueberry$Number.of.visits, simplify = "params")

#Plot
plot(Rabbiteye.Blueberry$PollinationValue ~ jitter(Rabbiteye.Blueberry$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(Rabbiteye.Blueberry$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")

rbb_data <- data.frame(crop = c("Rabbiteye.Blueberry"),
                        D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                        D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls, 
                        a = mf["a"],
                        b = mf["b"],
                        c = mf["c"],
                        OVR = calculate_required_visits(a = mf["a"],
                                                        b = mf["b"],
                                                        c_val = mf["c"]))
all_data <- rbind(all_data, rbb_data)
```

Quite similar fits again...

## Watermelon

```{r}
#head(Watermelon)
#unique(Watermelon$Pollinator) 
#unique(Watermelon$PollinationMetric) 

#fit models
model_nls <- fit_data(seedset = Watermelon$PollinationValue, 
                      visitation = Watermelon$Number.of.visits, simplify = "model", a_start = 75, b_start = 300)
model_lm <- lm(Watermelon$PollinationValue ~ Watermelon$Number.of.visits)
model_cuadratic <- lm(PollinationValue ~ Number.of.visits + I(Number.of.visits^2), data = Watermelon)

#Test fit
AIC_nls <- AIC(model_nls) #1188.206
AIC_lm <- AIC(model_lm) #1195.297
AIC_cuadratic <- AIC(model_cuadratic) #1188.724

#Extract R2
R2_nls <- pseudeR2(model_nls, Watermelon$PollinationValue) #0.1
R2_lm <- summary(model_lm)$r.squared #0.09 #again best!
R2_cuadratic <- summary(model_cuadratic)$r.squared #0.1

#Extract a, b and c values
mf <- fit_data(seedset = Watermelon$PollinationValue, 
               visitation = Watermelon$Number.of.visits, simplify = "params", 
               a_start = 75, b_start = 300)

#Plot
plot(Watermelon$PollinationValue ~ jitter(Watermelon$Number.of.visits), 
     las = 1, xlab = "Visits", ylab = "Fruit weigth", col = "grey")
plot_visits(mf["a"], mf["b"], mf["c"], to_ = max(Watermelon$Number.of.visits), 
            add_ = TRUE, col = "#E69F00")
abline(model_lm, col = "#56B4E9")
curve(predict(model_cuadratic, newdata = data.frame(Number.of.visits = x)), 
      add = TRUE, col = "#009E73")
wm_data <- data.frame(crop = c("Watermelon"),
                      D_R2 = max(R2_lm, R2_cuadratic)- R2_nls,
                      D_AIC = max(AIC_lm, AIC_cuadratic)- AIC_nls,
                        a = mf["a"],
                        b = mf["b"],
                        c = mf["c"],
                        OVR = calculate_required_visits(a = mf["a"],
                                                        b = mf["b"],
                                                        c_val = mf["c"]))
all_data <- rbind(all_data, wm_data)

#save dataset
write.csv(all_data, "data/parameters_real_crops.csv")
```

# Now let's run simulations for this same crops

We use paremetres from the litearture.

```{r}
#load params
d <- read.csv("data/VisitationRates_ParameterData.csv")
head(d)

#param a is in %
d$a <- (d$a/ d$b)*100

#SV we can take the average
d$SV <- rowMeans(d[,c("Honeybee", "Bumblebee", 
                  "Other.wild.bee", "Hoverfly", "Other")], na.rm = TRUE)
#
d$SV[which(d$Crop == "strawberry")] <- 150 #Using max efficiency...

#WATERMELON
#Now we can calculate the curve (c):
params <- calculate_visits(a = d$a[1], b = d$b[1], SVD = d$SV[1], 
                           plot_ = TRUE, to_ = 30)
#warning can be dissmissed.

#And visits needed.
OVR_model <- calculate_required_visits(a = d$a[1], b = d$b[1], c_val = params$c_parameter) 

#this is going to look baaad. The issue is that the literature data indicates efficiencies of ~200 seeds per single visit, while the empirical data suggest ~50. When using  "SVD = 50", the number of visits is 22, much closer to the reality.

#store data along with the other data
#load data
all_data <- read.csv("data/parameters_real_crops.csv")
all_data$a_model <- NA
all_data$b_model <- NA
all_data$c_model <- NA
all_data$OVR_model <- NA
#add
all_data$a_model[7] <- d$a[1]
all_data$b_model[7] <- d$b[1]
all_data$c_model[7] <- params$c_parameter
all_data$OVR_model[7] <- OVR_model

#loop for all? As only 5, I'll do manually
#highbush blueberry
#Now we can calculate the curve (c):
params <- calculate_visits(a = d$a[2], b = d$b[2], SVD = d$SV[2], 
                           plot_ = TRUE, to_ = 30)
#warning can be dissmissed.
#And visits needed.
OVR_model <- calculate_required_visits(a = d$a[2], b = d$b[2], c_val = params$c_parameter) 
#store data along with the other data
all_data$a_model[4] <- d$a[2]
all_data$b_model[4] <- d$b[2]
all_data$c_model[4] <- params$c_parameter
all_data$OVR_model[4] <- OVR_model

#apple
#Now we can calculate the curve (c):
params <- calculate_visits(a = d$a[3], b = d$b[3], SVD = d$SV[3], 
                           plot_ = TRUE, to_ = 30)
#warning can be dissmissed.
#And visits needed.
OVR_model <- calculate_required_visits(a = d$a[3], b = d$b[3], c_val = params$c_parameter) 
#store data along with the other data
all_data$a_model[1] <- d$a[3]
all_data$b_model[1] <- d$b[3]
all_data$c_model[1] <- params$c_parameter
all_data$OVR_model[1] <- OVR_model

#strawberry
#Now we can calculate the curve (c):
params <- calculate_visits(a = d$a[5], b = d$b[5], SVD = d$SV[5], 
                           plot_ = TRUE, to_ = 30)
#warning can be dissmissed.
#And visits needed.
OVR_model <- calculate_required_visits(a = d$a[5], b = d$b[5], c_val = params$c_parameter) 
#store data along with the other data
all_data$a_model[2] <- d$a[5]
all_data$b_model[2] <- d$b[5]
all_data$c_model[2] <- params$c_parameter
all_data$OVR_model[2] <- OVR_model

#lima bean
#Now we can calculate the curve (c):
params <- calculate_visits(a = d$a[6], b = d$b[6], SVD = d$SV[6], 
                           plot_ = TRUE, to_ = 30)
#warning can be dissmissed.
#And visits needed.
OVR_model <- calculate_required_visits(a = d$a[6], b = d$b[6], c_val = params$c_parameter) 
#store data along with the other data
all_data$a_model[3] <- d$a[6]
all_data$b_model[3] <- d$b[6]
all_data$c_model[3] <- params$c_parameter
all_data$OVR_model[3] <- OVR_model

#save file
write.csv(all_data, file = "data/parameters_real_crops.csv")

#plot 1:1 plot
all_data <- all_data[-4,]
plot(all_data$OVR, all_data$OVR_model, ylim = c(0,100), xlim = c(0,100), 
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)
#Not too satisfactory... specially strawberry, but also watermelon and lima bean
plot(all_data$a, all_data$a_model, ylim = c(0,50), xlim = c(0,50),
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)

plot(all_data$b, all_data$b_model, ylim = c(0,450), xlim = c(0,450),
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)

plot(all_data$c, all_data$c_model, ylim = c(0,4), xlim = c(0,4),
     col = c("green", "red", "brown", "blue","darkgreen"), las = 1)
abline(0, 1)
```

# Finally let's run it for all crops

```{r}
#re-read data
dat <- read.csv("data/VisitationRates_ParameterData.csv")
head(dat)
dat <- dat[-24,] #removing cheak peas, which is wierd as a = b = SV

#param a is in %
dat$a <- (dat$a / dat$b)*100

#SV we can take the average
dat$SV <- rowMeans(dat[,c("Honeybee", "Bumblebee", 
                      "Other.wild.bee", "Hoverfly", "Other")], na.rm = TRUE)
dat$SV[which(dat$Crop == "cherry")] <- 0.25 #Using only wild bees.
dat$SV[which(dat$Crop == "strawberry")] <- 150 #Using max efficiency...

#output data frame
out <- data.frame(crop = rep(NA,nrow(dat)), 
                  family = rep(NA,nrow(dat)),
                  a = rep(NA,nrow(dat)), 
                  b = rep(NA,nrow(dat)), 
                  c = rep(NA,nrow(dat)), 
                  SVE = rep(NA,nrow(dat)), 
                  OVR = rep(NA,nrow(dat)))
#loop it
for(i in 1:nrow(dat)){
  #extract c
  params <- calculate_visits(a = dat$a[i], b = dat$b[i], SVD = dat$SV[i], 
                             plot_ = FALSE)
  #warning can be dissmissed.
  #And visits needed.
  OVR_model <- calculate_required_visits(a = dat$a[i], b = dat$b[i], c_val = params$c_parameter) 
  #store the data
  out$crop[i] <- dat$Crop[i]
  out$family[i] <- dat$family[i]
  out$a[i] <- dat$a[i]
  out$b[i] <- dat$b[i]
  out$c[i] <- params$c_parameter
  out$SVE[i] <- dat$SV[i]
  out$OVR[i] <- OVR_model
}

bymedian <- with(out, reorder(family, -OVR, median))
boxplot(out$OVR ~ bymedian, las = 2, xlab = "")
bymedian <- with(out, reorder(crop, -OVR, median))
boxplot(out$OVR ~ bymedian, las = 2, xlab = "")

#parameter influence?
scatter.smooth(out$OVR ~ out$a, las = 1)
scatter.smooth(out$OVR ~ out$b, las = 1)
scatter.smooth(out$OVR ~ out$c, las = 1)
scatter.smooth(out$OVR ~ out$SVE, las = 1)
#It's a combination, but c is equivalent to OVR 

```
